---
title: "Coursera Project"
author: "Atiq ur Rehman"
date: "7/18/2020"
output: html_document
---

##Assignment Instructions
#1.Code for reading in the dataset and/or processing the data
#2.Histogram of the total number of steps taken each day
#3.Mean and median number of steps taken each day
#4.Time series plot of the average number of steps taken
#5.The 5-minute interval that, on average, contains the maximum number of steps
#6.Code to describe and show a strategy for imputing missing data
#7.Histogram of the total number of steps taken each day after missing values are imputed
#8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
#9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


```{r, echo=TRUE}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
  filegiven <- tempfile()
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",filegiven)
  unzip(filegiven)
  unlink(filegiven)
}

activity<-read.csv("activity.csv")
```

``` {r, echo = TRUE}
dim(activity)
names(activity)
head(activity)
str(activity)
```

``` {r, echo = TRUE}
stepsperday <-aggregate(steps~date, activity, sum)
png("plot1.png")

hist(stepsperday$steps, xlab = "Class", ylab = "Number of days", main = "Total Number of Steps taken each day" )

dev.off()
```

``` {r, echo = TRUE}
mean(stepsperday$steps)
median(stepsperday$steps)

```

``` {r, echo = TRUE}

meanstepsbyinterval <- aggregate(steps~interval, activity, mean)
with(meanstepsbyinterval, plot(interval, steps, type = "l"))
png("plot2.png")
dev.off()
```

``` {r, echo = TRUE}
meanstepsbyinterval[which.max(meanstepsbyinterval[,2]),1]

```

``` {r, echo = TRUE}
missingIndex<-is.na(activity[,1])
m <- mean(meanstepsbyinterval$steps)
activity1 <- activity
activity1[missingIndex,1]<-m
head(activity1)

```

``` {r, echo = TRUE}
stepsperday1 <-aggregate(steps~date, activity1, sum)
hist(stepsperday1$steps, xlab="Class", ylab="Number of Days", main="Number of Steps taken each day after missing values are imputed")
png("plot3.png")
dev.off()

```

``` {r, echo = TRUE}
stepsperday1 <-aggregate(steps~date, activity1, sum)

mean(stepsperday1$steps)
median(stepsperday1$steps)

```

``` {r, echo = TRUE}
meanstepsbyinterval <- aggregate(steps~interval, activity, mean)
with(meanstepsbyinterval, plot(interval, steps, type = "l"))

meanstepsbyinterval[which.max(meanstepsbyinterval[,2]),1]

png("plot4.png")

dev.off()

```

``` {r, echo = TRUE}
activity1$date<-as.Date(activity1$date)
library(dplyr)

activity2 <- activity1 %>%
  mutate(dayType= ifelse(weekdays(activity1$date)=="Saturday" | weekdays(activity1$date)=="Sunday", "Weekend", "Weekday"))

head(activity2)

```

``` {r, echo = TRUE}
meanstespbydaytypeandinterval <- activity2 %>%
  group_by(dayType, interval) %>%
  summarize(stepsperday=sum(steps))

head(meanstespbydaytypeandinterval)

library(lattice)
with(meanstespbydaytypeandinterval, 
     xyplot(stepsperday ~ interval | dayType, 
            type = "l",      
            main = "Total Number of Steps within Intervals by dayType",
            xlab = "Daily Intervals",
            ylab = "Average Number of Steps"))

png("plot5.png")
dev.off()

```